---
title: "Data Assembly"
output: html_document
editor_options: 
chunk_output_type: inline
---
PRIMARY PURPOSE: 
The purpose of this .Rmd is to assemble the data specific to this research into a training and test set that can be used with minial adjustment and cleaning for the analysis of this study. The pre-processed data for this study was downloaded from the NHANES website (https://wwwn.cdc.gov/nchs/nhanes/NhanesCitation.aspx), and are required to run this file. However, these files should be provided with a copy of the .Rmd. It is important to note that this file will only run if output is set to in-line. Additionally, the file directory should be changed to match the location of the necessary NHANES data. 


```{r setup, include=FALSE}
#### Attaching appropriate packages 
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) ###plotting 
library(dplyr) ### data manipulation 
library(readr) ### file reading 
library(Hmisc) ### additional features 
#### Setting working directory. CHANGE TO MATCH LOCATION OF NECESSARY NHANES DATA 
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/buehl/Documents/classes/senior-research"))
```



```{r}
#___________________________________________________________________________________________________
#### READING DATA FROM FOLDER 
#### reading in all .XPT SAS files from NHANES_DATA folder and labeling them accordingly. Each of these files contains variables relevant to this study and come from various portions of the NHANES survey. 
#___________________________________________________________________________________________________

nhanes_file_names <- list.files(path = "NHANES_DATA")
for(i in 1:length(nhanes_file_names))
{
  assign(paste(nhanes_file_names[i]), sasxport.get(paste0("NHANES_DATA/", nhanes_file_names[i])))
}
```

```{r}
#___________________________________________________________________________________________________
#### CONSOLIDATING COHORT YEARS FOR EACH DATA FRAME 
#### The data frames containing relevant variables are split up by their respective cohort year, so they need to be horizontally merged. This code works as NHANES variables are consists between years with the exception of the cholesterol data. The necessary adjustments have been made below. 

#### This code should be generalizable, but for some reason wouldn't work. I posted about this issue on Stack OVerflow (https://stackoverflow.com/questions/55151729/having-trouble-generalizing-a-data-cleaning-solution-to-a-function-in-r). 
#___________________________________________________________________________________________________


years <- c("_1999.XPT","_2003.XPT","_2005.XPT","_2007.XPT","_2009.XPT","_2011.XPT","_2013.XPT","_2015.XPT")

####Blood Pressure 
bloodpressure <- bloodpressure_1999.XPT
for(i in 2:length(years))
{
  bloodpressure <- bind_rows(bloodpressure,get(paste0("bloodpressure",years[i])))    
}

####Bad Cholesterol
bad_cholesterol <- bad_cholesterol_1999.XPT
for(i in 2:length(years))
{
  bad_cholesterol <- bind_rows(bad_cholesterol,get(paste0("bad_cholesterol",years[i])))    
}

####body measure 
body_measure <- body_measure_1999.XPT
for(i in 2:length(years))
{
  body_measure <- bind_rows(body_measure,get(paste0("body_measure",years[i])))    
}

####demographic 
demographic <- demographic_1999.XPT
for(i in 2:length(years))
{
  demographic <- bind_rows(demographic,get(paste0("demographic",years[i])))    
}


####diabetes 
diabetes <- diabetes_1999.XPT
for(i in 2:length(years))
{
  diabetes <- bind_rows(diabetes,get(paste0("diabetes",years[i])))    
}


####glucose 
glucose <- glucose_1999.XPT
for(i in 2:length(years))
{
  glucose <- bind_rows(glucose,get(paste0("glucose",years[i])))    
}



####diabetes 
diabetes <- diabetes_1999.XPT
for(i in 2:length(years))
{
  diabetes <- bind_rows(diabetes,get(paste0("diabetes",years[i])))    
}


####hypertension
hypertension <- hypertension_1999.XPT
for(i in 2:length(years))
{
  hypertension <- bind_rows(hypertension,get(paste0("hypertension",years[i])))    
}


####total and good cholesterol 
#### Making adjustment for cholesterol 
cholesterol <- merge(get(paste0("cholesterol",years[3])),get(paste0("good_cholesterol",years[3])))
for(i in 4:length(years))
{
  
  a <- merge(get(paste0("cholesterol",years[i])),get(paste0("good_cholesterol",years[i])))
  cholesterol <- bind_rows(cholesterol,a)
}
total_good_cholesterol_1999.XPT
cholesterol <- bind_rows(cholesterol, total_good_cholesterol_1999.XPT)
cholesterol <- bind_rows(cholesterol, total_good_cholesterol_2001.XPT)
cholesterol <- bind_rows(cholesterol, total_good_cholesterol_2003.XPT)
```

```{r}
#___________________________________________________________________________________________________
#### EXTRACTING IMPORTANT VARIABLES FROM DATA FRAMES AND CREATING TRAINING AND TEST 
#### Not all of the variables from the dataframes are required for the study and thus are filtered out below. Then, the variables are combined into a training and test set. Additionally, OUTLIERS FOR BLOOD PRESSURE are removed here.  
#___________________________________________________________________________________________________

#___________________________________________________________________________________________________
#### DV: Blood Pressure 
  ### selecting blood relevant blood pressures 
  dv_bp <- bloodpressure %>% select(seqn,bpxsy1,bpxsy2,bpxsy3,bpxdi1,bpxdi2,bpxdi3)

  ### throwing out outliers
  dv_bp <- dv_bp %>% filter(bpxsy1 != 0 & bpxsy2 != 0 & bpxsy3 != 0 & bpxdi1 != 0 & bpxdi2 != 0 & bpxdi3 != 0 & bpxsy1 < 200 & bpxsy2 < 200 & bpxsy3 < 200 & bpxdi1 < 200 & bpxdi2,bpxdi3 < 200 & bpxdi1 > 27 & bpxdi2 > 27 & bpxdi3 > 27)
  summary(dv_bp)
  
  #### calculating mean 
  dv_bp$dia_mean <- rowMeans(dv_bp[,c("bpxdi1","bpxdi2","bpxdi3")])
  dv_bp$sys_mean <- rowMeans(dv_bp[,c("bpxsy1","bpxsy2","bpxsy3")])
#___________________________________________________________________________________________________
  
#___________________________________________________________________________________________________
#### IV: BMI Info 
bmi_variables <- body_measure %>% select(seqn,bmxarmc, bmxarml, bmxleg, bmxrecum, bmxsad1, bmxht, bmxwaist, bmxwt, bmxbmi)

#### IV: Demographic 
demo_variables <- demographic %>% select(seqn,riagendr, ridageyr,ridreth3)

demo_variables$ridreth3 <- ifelse(demo_variables$ridreth3 <= 2, 1,ifelse(demo_variables$ridreth3 == 3,2,ifelse(demo_variables$ridreth3 == 4,3, ifelse(demo_variables$ridreth3 == 6, 4, ifelse(demo_variables$ridreth3 == 7,5,NA)))))

#### dumming out 
  #### race
  demo_variables <- demo_variables %>% mutate(hispanic = ifelse(ridreth3 == 1,1,0),white = ifelse(ridreth3 == 2,1,0),black = ifelse(ridreth3 == 3,1,0),asian = ifelse(ridreth3 == 4,1,0),mult = ifelse(ridreth3 == 5,1,0))
  
  #### gender 
  demo_variables$riagendr[demo_variables$riagendr == 2] <- 0  

#### IV: Diabetes 
dia_variables <- diabetes %>% select(seqn,diq010) %>% filter(diq010 <= 3)
summary(dia_variables)
dia_variables$diq010[dia_variables$diq010 == 3] <- 1  
dia_variables$diq010[dia_variables$diq010 == 1] <- 1
dia_variables$diq010[dia_variables$diq010 == 2] <- 0

#### IV: Cholesteral 
chol_variables <- cholesterol %>% select(seqn,lbdhdd,lbxtc)
bad_chol_variables <- bad_cholesterol %>% select(seqn,lbdldl)



#### IV: Hypertension Medication
hypertension_variables <- hypertension %>% select(seqn,bpq020,bpq040a)
  #### only yes and nos 
  hypertension_variables <- hypertension_variables %>% filter(bpq020 == 1 | bpq020 == 2)
  hypertension_variables$bpq020[hypertension_variables$bpq020 == 2] <- 0
  
  #### nas are nos for hypertension 
  hypertension_variables$bpq040a[is.na(hypertension_variables$bpq040a)] <- 0
  
  #### dumming out hypertension
  hypertension_variables$bpq040a <- ifelse(hypertension_variables$bpq040a == 0, 0, ifelse(hypertension_variables$bpq040a == 1, 1, ifelse(hypertension_variables$bpq040a == 2, 0, ifelse(hypertension_variables$bpq040a == 7,NA, ifelse(hypertension_variables$smq040 == 9,NA,NA)))))

####IV: glucose 
glu_variables <- glucose %>% select(seqn,lbxglu)
#___________________________________________________________________________________________________
#___________________________________________________________________________________________________

#### merging into single grame 
lin_frame <- dv_bp 
variables <- c("bmi_variables","demo_variables","dia_variables","bad_chol_variables","chol_variables","glu_variables","hypertension_variables")
for(i in 1:length(variables))
{
  lin_frame <- merge(lin_frame,eval(as.name(variables[i])),by=c("seqn"))
}
summary(lin_frame) #### summary to check everything looks reasonable 

#### 75% of the sample size
smp_size <- floor(0.75 * nrow(lin_frame))

#### set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(lin_frame)), size = smp_size)

#### Create training and test set 
lin_frame <- lin_frame[train_ind, ]
lin_frame_test <- lin_frame[-train_ind, ]

```

```{r}
#___________________________________________________________________________________________________
#### MISC CLEANING 
#### remove misc. variables and drop NAs 
#___________________________________________________________________________________________________

#### test frame 
lin_frame_test$bmxrecum <- NULL 
lin_frame_test <- subset(lin_frame_test, select=-c(seqn,bpxsy1,bpxsy2,bpxsy3,bpxdi1,bpxdi2,bpxdi3)) 
lin_frame_test <- na.omit(lin_frame_test)
lin_frame_test$id <- seq.int(nrow(lin_frame_test))
####Prediction Frame 
lin_frame$bmxrecum <- NULL 
lin_frame_clean <- subset(lin_frame, select=-c(seqn,bpxsy1,bpxsy2,bpxsy3,bpxdi1,bpxdi2,bpxdi3)) 
lin_frame_clean <- na.omit(lin_frame_clean)



#### Blood Pressure Frame 
lin_frame_bp <- na.omit(lin_frame)
summary(lin_frame_bp)


```