---
title: "linear_model"
output: html_document
editor_options: 
chunk_output_type: console
---
PRIMARY PURPOSE: 
The purpose of this .Rmd is to build a linear model for both systolic and diastolic blood pressure to establish a reasonable baseline for the study. The mse data frame created in this file is required in all subsquent modelling files and thus it is essential that this file is run before other modelling files. The mse data frame stores information about the performance of the various models in this study. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
library(matrixStats)
library(reshape2)
library(car) 
library(MASS)
library(jtools) ###exporting linear models 
```



```{r}
#___________________________________________________________________________________________________
#### CONSTRUCTING LINEAR MODEL FOR SYSTOLIC BLOOD PRESSURE 
#### An initial model is fit to the data in model 1. Assumptions are then amended and a better model is fit in model 2. Lastly, model 2's performance on the test set is recoreded in the mse data frame. The mse data frame will keep predicion information from models in subsquent files and thus is essential. 
#___________________________________________________________________________________________________

#### Systolic prediction 
#### correlation
sys_lin <- subset(lin_frame_clean, select=-c(dia_mean,ridreth3,mult))

#### model 1: linear regression
bpsys_linear1 = lm(sys_mean~.,sys_lin)
summary(bpsys_linear1)

  #### looking at variance explained by interactions
  bpsys_interact = lm(sys_mean~(.)^2,sys_lin)
  anova(bpsys_interact)

  #### residual assumption 
  #### overal residual values 
  plot(bpsys_linear1$fitted.values,bpsys_linear1$residuals)
  
  #### all residuals 
  crPlots(bpsys_linear1)

  
  #### qq-plot
  qqPlot(bpsys_linear1,ylab="Standardized Residuals", xlab="Normal Scores")
  
  #### correcting bmxsad1
  res_correct = lm(sys_mean~log(bmxsad1),sys_lin)
  res_not = lm(sys_mean~(bmxsad1),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  #### correcting bmxwt
  res_correct = lm(sys_mean~log(bmxwt),sys_lin)
  res_not = lm(sys_mean~(bmxwt),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  
  #### correcting year 
  res_correct = lm(sys_mean~I(ridageyr^2),sys_lin)
  res_not = lm(sys_mean~(ridageyr),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  ####correcting lbdldl
  res_correct = lm(sys_mean~log(lbdldl),sys_lin)
  res_not = lm(sys_mean~(lbdldl),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)

  
  ####correcting lbxtc
  res_correct = lm(sys_mean~log(lbxtc),sys_lin)
  res_not = lm(sys_mean~(lbxtc),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  ####correcting lbxglu
  res_correct = lm(sys_mean~log(lbxglu),sys_lin)
  res_not = lm(sys_mean~(lbxglu),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)

#### model 2: asumption amended linear regression
bpsys_linear2 = lm(sys_mean~ bmxarmc + bmxsad1 + bmxht + bmxwaist + bmxwt + riagendr + ridageyr + I(ridageyr^2) + hispanic + white + black + diq010 + lbdldl + log(lbxtc) + log(lbxglu) + bpq020 + bpq040a + bmxarmc:ridageyr + riagendr:lbdldl + riagendr:lbxglu + bmxsad1:diq010 + riagendr:ridageyr + bmxsad1:bpq040a  + ridageyr:white + ridageyr:diq010 + ridageyr:lbdldl  , sys_lin)
summary(bpsys_linear2)


####keep in original for interaction and transformation 
  #### residual assumption 
  #### overal residual values 
  plot(bpsys_linear2$fitted.values,bpsys_linear2$residuals)
  
  #### all residuals 
  crPlots(bpsys_linear2)

  
  #### qq-plot
  qqPlot(bpsys_linear2,ylab="Standardized Residuals", xlab="Normal Scores")



#### prediction 
sys_predictions = predict(bpsys_linear2, newdata = lin_frame_test)
summary(sys_predictions)
summary(lin_frame_test$sys_mean)

### MSE data frame 
mse <- data.frame(id = seq.int(nrow(lin_frame_test)), true_sys = lin_frame_test$sys_mean, true_dia = lin_frame_test$dia_mean, sys_lin_mse = (lin_frame_test$sys_mean - sys_predictions)^2, sys_lin = sys_predictions) 

```

```{r}
#___________________________________________________________________________________________________
#### CONSTRUCTING LINEAR MODEL FOR DIASTOLIC BLOOD PRESSURE 
#### An initial model is fit to the data in model 1. Assumptions are then amended and a better model is fit in model 2. Lastly, the model 2's performance on the test set is added to the mse data frame 
#___________________________________________________________________________________________________

#### Diastolic prediction 
#### correlation
dia_lin <- subset(lin_frame_clean, select=-c(sys_mean,ridreth3,mult))
cor(dia_lin)


#### model 1: linear regression
bpdia_linear1 = lm(dia_mean~.,dia_lin)
summary(bpdia_linear1)

  #### looking at variance explained by interactions
  bpsys_interact = lm(dia_mean~(.)^2,dia_lin)
  anova(bpsys_interact)

  #### residual assumption 
  #### overal residual values 
  plot(bpdia_linear1$fitted.values,bpdia_linear1$residuals)
  
  #### all residuals 
  crPlots(bpdia_linear1)

  
  #### qq-plot
  qqPlot(bpdia_linear1,ylab="Standardized Residuals", xlab="Normal Scores")
  
  #### correcting lbdldl
  res_correct = lm(sys_mean~log(lbdldl),sys_lin)
  res_not = lm(sys_mean~(lbdldl),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  #### correcting lbxglu
  res_correct = lm(sys_mean~log(lbxglu),sys_lin)
  res_not = lm(sys_mean~(lbxglu),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  #### correcting lbxtc
  res_correct = lm(sys_mean~log(lbxtc),sys_lin)
  res_not = lm(sys_mean~(lbxtc),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  #### correcting bmxwt
  res_correct = lm(sys_mean~log(bmxwt),sys_lin)
  res_not = lm(sys_mean~(bmxwt),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)
  
  #### correcting bmxbmi
  res_correct = lm(sys_mean~log(bmxbmi),sys_lin)
  res_not = lm(sys_mean~(bmxbmi),sys_lin)
  crPlots(res_not)
  crPlots(res_correct)

#### model 2: linear regression
bpdia_linear2 = lm(dia_mean~ bmxarmc + bmxsad1 + log(bmxwt) + riagendr + ridageyr + I(ridageyr^2) + white + black + asian + diq010 + log(lbdldl) + lbxtc + bpq020 + bpq040a + ridageyr:white +  bmxsad1:ridageyr + ridageyr:bpq040a + riagendr:lbdldl,dia_lin)
summary(bpdia_linear2)


  #### residual assumption 
  #### overal residual values 
  plot(bpdia_linear2$fitted.values,bpdia_linear2$residuals)
  
  #### all residuals 
  crPlots(bpdia_linear2)

  
  #### qq-plot
  qqPlot(bpdia_linear2,ylab="Standardized Residuals", xlab="Normal Scores")

#### prediction 
dia_predictions = predict(bpdia_linear2, newdata = lin_frame_test)
summary(dia_predictions)
summary(lin_frame_test$dia_mean)

#### MSE Frame update
mse$dia_lin_mse <- (lin_frame_test$dia_mean - dia_predictions)^2
mse$dia_lin  <- dia_predictions
mal_pred <- cbind(mse$dia_lin,mse$sys_lin)
mal_true <- cbind(mse$true_dia,mse$true_sys)
mse$lin_mal <- mahalanobis(mal_pred, mal_true, cov(mal_pred,mal_true))


```

```{r}
#___________________________________________________________________________________________________
#### COMPARING PREDICTED VALEUS TO ACTUAL VALUES FROM LINEAR MODELS 
#___________________________________________________________________________________________________
ggplot(mse, aes(true_dia,dia_lin))+
    geom_point() + geom_abline(aes(intercept=0, slope=1, col = "slope = 1")) + geom_smooth(method="lm")  + labs( 
   x="Predicted Diastolic", y="True Diastolic") + theme_minimal() 

ggplot(mse, aes(true_sys,sys_lin))+
    geom_point() + geom_abline(aes(intercept=0, slope=1, col = "slope = 1")) + geom_smooth(method="lm") + labs( 
   x="Predicted Systolic", y="True Systolic") + theme_minimal()


```


```{r}
#___________________________________________________________________________________________________
#### Exporting Linear MOdels 
#___________________________________________________________________________________________________
table1 <- export_summs(bpsys_linear2,bpdia_linear2, scale = TRUE)
write.csv(table1, "model.csv")
```


